Project1
========================================================
Read table into R and change into data table
```{r}
activity <- read.csv("C:/Users/David/Dropbox/Coursea/Reproducible Research/repdata-data-activity/activity.csv",na.strings = "NA")
library("data.table")
actDT=as.data.table(activity)
```

Get a summary and structure of the dataset
```{r}
summary(actDT)
str(actDT)
```
                 
Change date to Date field
```{r}
actDT[,date:=as.Date(date,"%Y-%m-%d")]
```

##What is the mean total number of steps taken per day?

Calculate the total number of steps take each day
```{r}
total.steps.by.date=actDT[,sum(steps,na.rm=T),by=date]
```

Change column names to dates and total.steps
```{r}
setnames(total.steps.by.date,1:2,c("dates","total.steps"))
```

Make a histogram of the total number of steps taken each day
```{r}
hist(total.steps.by.date[,total.steps],main="Histogram of the total number of steps taken each day",xlab="Total number of steps taken each day")
```

Calculate the mean and median total number of steps taken per day
```{r}
total.steps.by.date[,c(mean=mean(total.steps,na.rm=T),median=median(total.steps,na.rm=T))]

```

##What is the average daily activity pattern?

Calculate the the average number of steps taken all days by time interval
```{r}
avg.steps.by.interval=actDT[,mean(steps,na.rm=T),by=interval]
setnames(avg.steps.by.interval,1:2,c("interval","avg.steps"))
```

Make the time series plot 
```{r}
plot(avg.steps.by.interval$interval,avg.steps.by.interval$avg.steps,type="l",xlab="5-minute interval",ylab=" Average number of steps taken across all days",main="Average daily activity pattern")
```

calculate Which 5-minute interval, on average across all the days in the dataset,contains the maximum number of steps
```{r}
avg.steps.by.interval[avg.steps==max(avg.steps,na.rm=T),]
```

##Imputing missing values

Calculate and report the total number of missing values in the dataset 
```{r}
sum(!complete.cases(actDT))
```

fill in all missing values by the mean for that 5-minute interval
set interval as key column and merge actDT with avg.steps.by.interval
```{r}
setkey(actDT,interval)
setkey(avg.steps.by.interval,interval)
actDT=actDT[avg.steps.by.interval]
```

change steps from int to double before filling the NA with avg.steps
```{r}
actDT[,steps:=as.numeric(steps)]
actDT[is.na(steps),steps:=avg.steps]
```

number of missing steps
```{r}
actDT[,sum(is.na(steps))]
```

Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
actDT.filled=actDT[,1:3,with=F]
```

Calculate the total number of steps take each day for actDT.filled
```{r}
total.steps.by.date=actDT.filled[,sum(steps,na.rm=T),by=date]
```

change column names to dates and total.steps
```{r}
setnames(total.steps.by.date,1:2,c("dates","total.steps"))
```

Make a histogram of the total number of steps taken each day
```{r}
hist(total.steps.by.date[,total.steps],main="Histogram of the total number of steps taken each day with filled NA",xlab="Total number of steps taken each day")
```

Calculate the mean and median total number of steps taken per day
```{r}
total.steps.by.date[,c(mean=mean(total.steps,na.rm=T),median=median(total.steps,na.rm=T))]
```

**Do these values differ from the estimates from the first part of the assignment?** 
Yes

**What is the impact of imputing missing data on the estimates of the total daily number of steps?**
The mean and the median is now the same.

##Are there differences in activity patterns between weekdays and weekends?
```{r}
actDT.filled=actDT.filled[(weekdays(date)=="Saturday"|weekdays(date)=="Sunday"),weekday:="weekend"]
actDT.filled=actDT.filled[!(weekdays(date)=="Saturday"|weekdays(date)=="Sunday"),weekday:="weekday"]
actDT.filled=actDT.filled[,weekday:=factor(weekday)]
```

calculate the the average number of steps taken all days by time interval
```{r}
avg.steps.by.weekday=actDT.filled[,mean(steps),by=list(weekday,interval)]
setnames(avg.steps.by.weekday,1:3,c("weekday","interval","avg.steps"))
```

Make the time series plot 
```{r}
library("ggplot2")
qplot(interval,avg.steps,data=avg.steps.by.weekday,facets=weekday~.,geom="line")+labs(y="Number of steps")
```
